Concurrent Programming with CXL

### 2020-06-14 book/paper.tex
Adding description of machine instructions

\texttt{release\_wlock} executes \texttt{notifyAll \&(wcond)} as well as
\texttt{notify \&(wcond)}.
Again, because we do not keep track of the number of waiting readers or writers, we
have to conservatively assume that all waiting readers can enter, or, alternatively,
up to one waiting writer can enter.  So \texttt{release\_wlock} wakes up all
potential candidates.
There are two things to note here.  First, unlike split binary semaphores or Hoare
monitors, where multiple waiting readers would have to be signaled one at a time in a
baton-passing fashion (see Figure~\ref{fig:splitsema}), with Mesa monitors
all readers are awaked in one fell swoop using \texttt{notifyAll}.
Second, both readers and writers are awakened---this is ok because both execute
\texttt{wait} within a \texttt{while} loop, re-checking the condition that they
are waiting for.  So if both type of processes are waiting, either all the readers
get to enter next or one of the writers gets to enter next.

@@

>>>>

\texttt{release\_wlock} executes \texttt{notifyAll \&(wcond)} as well as
\texttt{notify \&(wcond)}.
Again, because we do not keep track of the number of waiting readers or writers, we
have to conservatively assume that all waiting readers can enter, or, alternatively,
up to one waiting writer can enter.  So \texttt{release\_wlock} wakes up all
potential candidates.
There are two things to note here.  First, unlike split binary semaphores or Hoare
monitors, where multiple waiting readers would have to be signaled one at a time in a
baton-passing fashion (see Figure~\ref{fig:RWsplitsema}), with Mesa monitors
all readers are awaked in one fell swoop using \texttt{notifyAll}.
Second, both readers and writers are awakened---this is ok because both execute
\texttt{wait} within a \texttt{while} loop, re-checking the condition that they
are waiting for.  So if both type of processes are waiting, either all the readers
get to enter next or one of the writers gets to enter next.

@@

\chapter{List of CXL Machine Instructions}

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
Address & compute address \\
\hline
Apply & pop $m$ and $i$ and apply $i$ to $m$, pushing a value \\
\hline
Assert & pop $b$ and check that it is \texttt{True} \\
\hline
AtomicDec & decrement the atomic counter of this context \\
\hline
AtomicInc & increment the atomic counter of this context \\
\hline
Choose & choose an element from the set on top of the stack \\
\hline
Del & delete a shared variable \\
\hline
DelVar $v$ & delete process variable $v$ \\
\hline
Dict & pop $n$ and $n$ key/value pairs and push a dictionary \\
\hline
Dup & duplicate the top element of the stack \\
\hline
Frame $m$ $a$ $r$& start method $m$ with arguments $a$,
initializing variables.  Last instruction is at $r$ \\
\hline
Go & restore context \\
\hline
JumpCond $e$ $p$ & pop expression and if equal to $e$ set program counter to $p$ \\
\hline
Jump $p$ & set program counter to $p$ \\
\hline
Load & load a shared variable and push it onto the stack \\
\hline
LoadVar $v$ & load process variable $v$ and push it onto the stack \\
\hline
Nary \textit{op} $n$ & apply $n$-ary operator \textit{op} to the top $n$ elements on the stack \\
\hline
Pop & pop a value of the stack and discard it \\
\hline
Push $c$ & push constant $c$ onto the stack \\
\hline
PushAddress $v$ & push the address of shared variable $v$ \\
\hline
Return & pop return address, push \texttt{result}, and restore program counter \\
\hline
Set & pop $n$ and $n$ elements and push a set of the elements \\
\hline
Spawn & pop tag, argument, and method and spawn a new context \\
\hline
Split & split a non-empty set in its minimum element and its remainder \\
\hline
Stop & save context \\
\hline
Store & pop a value from the stack and store it in a shared variable \\
\hline
StoreVar $v$ & pop a value from the stack and store it in process variable $v$ \\
\hline
\end{tabular}
